Positive Correlation Between Racialization, Gendering and Police Strip Searches in the Toronto Municipal, 2020-2021

### 2024-01-26 outputs/paper/paper.qmd
fine-tune writing and finalize paper

It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd]. Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021. Nonetheless, it's still worth looking at due the precedents of police power abuse [@covidStripSearch].

@@

Most relevant to the paper are the variables on the racial and gender identities of people who were arrested and then strip-searched. There are 9 unique racial groups for the `race` variables and 3 unique gender groups for the `gender` variables. To streamline this data, the names within the original dataset are simplified and then the dataset is split into a smaller dataset containing relevant columns ([see @tbl-cleaned_race_gender_data]). This dataset entails demographic and strip search-relevant information for each police arrest event, such as race, gender, whether the event involves a strip search, and if it is a strip search, are there any items found.

Specifically, the column 'Strip Search' signifies whether an arrest involves strip search. As for the 'Items Found' column, either a `1` or a `0` will appear in a data cell if the 'Strip Search' value of an event is a `1`. This is because the items found within a strip search is only available when a police arrest does involve strip searching.

@@

Nonetheless, the dataset only contains a binaristic view of gender, without regarding those who identify as non-binary or others, and essentially lumping all other gender identities as "Unidentified". This proves inconsistent against the constant public backlash agasint police for their power abuse of people belonging to the LGBTQI+ community in Toronto [@policeGender]. This point will be further discussed below.

@@

In general, there is a substantial 10% of strip searches being conducted without any reasons (see [@tbl-reasons-strip-search]). Commonly, strip searches are carried out by at least 1 or 2 reasons provided by the police, both at 29% (see [@tbl-reasons-strip-search]). As for items found during searches, at a whooping 63%, most strip search enforcement doesn't garner any suspected items (see [@tbl-items-strip-search]). This critically denotes how many strip searches aren't motivated by civic considerations but are moreso carried out by intuition, or personal biases that can extend to racial profiling. It signifies an alarming potential for misconducted strip searches that are racially or sexually motivated. To visualize this data, @fig-reasons-race and @fig-reasons-gender both map, respectively, the racial and gender distribution of strip searches by the total number of reasons each strip search has.

@@

Regarding @fig-reasons-race, due to the sheer number of police searches that people who are racially black and white have compared to other groups, this paper subsections the dataset on racialized police searches into two graphs. Both of these graphs are devoid of data from people are racially unknown or don't identify with any races--"None" or "Unknown or Legacy"--to ensure graph visibility. In specifics, graph a) of @fig-reasons-race contains only the distribution from people who racially identify as black and white, while graph b) of @fig-reasons-race contains the distribution of all other minority groups--except for black people--along with white people. This ensures data clarity between groups with higher counts of strip searches and groups with lower counts of strip searches. In addition, to ensure that data is proportionate to one another, the percentages for each racial group when subdivided into each of the different sums of reasons are calculated and used instead of `count`.

Based on part a) of @fig-reasons-race, the number of black people who are strip searched surpasses white people on all counts of total number of reasons, except for instances where the police only has one reason to conduct strip search. This again signifies an alarming rate of racialized strip searches against black people despite their small population make-up in Toronto (@ethnicCensus). For part b) of @fig-reasons-race, the percentages between other minority groups fluctuate. But intriguingly, compared to white people, all of other minority groups are more likely to be strip searched regardless of the number of reasons at hand. To note, at 11%, 11% and 10% each, people of "Middle-Eastern", "East/Southeast Asian" and "Indigenous" descent are most commonly strip searched without any provided reasons. "South Asian" in particular surpass all other racial groups when it comes to being strip searched because of 1 or 2 reasons for their actions at the time of arrest, at 34% and 35% respectively. This connotes that people of "South Asian" descent are more readily strip searched by police despite having a few reasons for their actions at the time of arrest. Speculatively, there could be a lot of colorism at play, considering how prerelevant racial profiling is to police power abuse, where South Asian and black people generally have a darker skin color compared to all of the other races mentioned [@skinColor].

>>>>

It's worth noting that this dataset may already contain much racial and gender bias, because it was retrieved from the very source that this paper is questioning: the Toronto Police Services [@stripSearchOpEd]. Then, the findings presented might not fully encapsulate the actual landscape of strip searching in Toronto from 2020 to 2021. Nonetheless, it's still worth looking at due to the precedents of police power abuse [@covidStripSearch].

@@

Most relevant to the paper are the variables on the racial and gender identities of people who were arrested and then strip-searched. After having their names simplified, this subset of data from the original dataset was generated, containing relevant columns on the social identities of people invovled in strip search ([see @tbl-cleaned_race_gender_data]).This dataset entails demographic and strip search-relevant information for each police arrest event, such as race, gender, whether the event involves a strip search, and any items found during serach. There are 9 unique racial groups for the `race` variables and 3 unique gender groups for the `gender` variables

Specifically, the column 'Strip Search' signifies whether an arrest involves strip searching. As for the 'Items Found' column, either a `1` or a `0` will appear in a data cell if the 'Strip Search' value of an event is a `1`. This is because the items found within a strip search is only available when a police arrest does involve strip searching.

@@

Nonetheless, the dataset only contains a binaristic view of gender, without regarding those who identify as non-binary or others, and essentially lumping all other gender identities as "Unidentified". This proves inconsistent against the constant public backlash agasint police for their power abuse of people belonging to the LGBTQI+ community in Toronto [@policeGender]; this point will be further discussed below.

@@

In general, there is a substantial 10% of strip searches being conducted without any reasons (see [@tbl-reasons-strip-search]). Commonly, strip searches are carried out by at least 1 or 2 reasons provided by the police, both at 29% (see [@tbl-reasons-strip-search]). As for items found during searches, at a whopping 63%, most strip search enforcement doesn't garner any suspected items (see [@tbl-items-strip-search]). This critically denotes how many strip searches aren't motivated by civic considerations but are moreso carried out by intuition, or personal biases that can extend to racial profiling. It signifies an alarming potential for misconducted strip searches that are racially or sexually motivated. To visualize this data, @fig-reasons-race and @fig-reasons-gender both map, respectively, the racial and gender distribution of strip searches by the total number of reasons each strip search has.

@@

Regarding @fig-reasons-race, due to the sheer number of police searches that people who are racially black and white have compared to other groups, this paper subsections the dataset on racialized police searches into two graphs. Both of these graphs are devoid of data from people are racially unknown or don't identify with any races--"None" or "Unknown or Legacy"--to ensure graph visibility. In specifics, graph (a) of @fig-reasons-race contains only the distribution from people who racially identify as black and white, while graph (b) of @fig-reasons-race contains the distribution of all other minority groups--except for black people--along with white people. This ensures data clarity between groups with higher counts of strip searches and groups with lower counts of strip searches. In addition, to ensure that data is proportionate to one another, the percentages for each racial group when subdivided into each of the different sums of reasons are calculated and used instead of `count`.

Based on part (a) of @fig-reasons-race, the number of black people who are strip searched surpasses white people on all counts of total number of reasons, except for instances where the police only has one reason to conduct strip search. This again signifies an alarming rate of racialized strip searches against black people despite their small population make-up in Toronto (@ethnicCensus). For part (b) of @fig-reasons-race, the percentages between other minority groups fluctuate. But intriguingly, compared to white people, all of other minority groups are more likely to be strip searched regardless of the number of reasons at hand. To note, at 11%, 11% and 10% each, people of "Middle-Eastern", "East/Southeast Asian" and "Indigenous" descent are most commonly strip searched without any provided reasons. "South Asian" in particular surpass all other racial groups when it comes to being strip searched because of 1 or 2 reasons for their actions at the time of arrest, at 34% and 35% respectively. This connotes that people of "South Asian" descent are more readily strip searched by police despite having a few reasons for their actions at the time of arrest. Speculatively, there could be a lot of colorism at play, considering how prerelevant racial profiling is to police power abuse, where South Asian and black people generally have a darker skin color compared to all of the other races mentioned [@skinColor].
